---
title: "Course Project 1"
author: "Diego Gajardo"
date: "19 de mayo de 2016"
output: html_document
---


###Loading and preprocesing the data

We read the data and remove the NAs cases:

```{r cache = TRUE}
rawdata <- read.csv("activity.csv")
ok <- complete.cases(rawdata)
cleandata <- rawdata[ok,]
cleandata$date <- factor(cleandata$date)
```

Then we take "interval" variable as factor to make future calculations:

```{r}
rawdata$interval <- as.factor(rawdata$interval)
cleandata$interval <- as.factor(cleandata$interval)
```


###What is mean total number of steps taken per day?

The total number of steps taken per day is given by:

```{r}
total_steps <- tapply(cleandata$steps,cleandata$date,sum)
(total_steps)
```

Now we make a histogram of the total steps:

```{r}
hist(total_steps, col = "green", xlab = "Total Steps", main = "Histogram of total steps per day")
```

Finally, we take the *mean* and *median* of the total number of steps:

```{r}
total_mean <- mean(total_steps)
total_median <- median(total_steps)
total_mean
total_median
```


###What is the average daily activity pattern?

First we take the average number of steps on each interval, and then we plot:

```{r}
avg <- tapply(cleandata$steps, cleandata$interval, mean)
plot(avg ~ levels(cleandata$interval), type = "l", col = "blue", xlab ="Intervals [minutes]", ylab = "Average of steps", main = "Average of steps per interval")
grid(nx = 50)
```

And the interval with the maximum number of steps:
```{r}
avg[avg == max(avg)]
```


###Imputing missing values

First we calculate the number of missing values:

```{r}
total_nas <- sum(!complete.cases(rawdata))
total_nas
```

Now we fill the misses values with the mean of the interval, and build a new dataset:

```{r}
na_cases <- which(!complete.cases(rawdata))
filldata <- rawdata
for (i in na_cases) {
        interval <- as.character(filldata[i,3])
        filldata[i,1] <- avg[interval]
}
head(filldata)
```

And we have a new dataset called *_filldata_*
Finally, we make an histogram of the total steps per day:

```{r}
total_steps_fill <- tapply(filldata$steps,filldata$date,sum)
hist(total_steps_fill, col = "red", xlab = "Total Steps", main = "Histogram of total steps per day within filled data")
```

And calculate mean and median for the new data:

```{r}
total_mean_fill <- mean(total_steps_fill)
total_median_fill <- median(total_steps_fill)
total_mean_fill
total_median_fill
```
It's easy to note that there is a little difference between the old and new median. The efect of filling the missing values is minimun, that is because we are using the average values to do it.


###Are there differences in activity patterns between weekdays and weekends?

First we create the new factor variable:

```{r}
filldata["day"] <- NA
for (i in seq(1,nrow(filldata))){
        filldata[i,"day"] <- ifelse(weekdays(as.Date(filldata[i,"date"])) == "sábado" | weekdays(as.Date(filldata[i,"date"])) == "domingo", "weekend","weekday")        
}
head(filldata)
```

And now we plot the average number of steps taken:

```{r}
library(ggplot2)
ggplot(filldata, aes(x = interval, y = steps)) + stat_summary(fun.y="mean", geom="line")+facet_wrap(~day)
```